Supplement to: On the propriety of restricted Boltzmann machines

### 2016-11-18 writing/appendix/appendix-center.Rmd
Nordman's edits

@@

{\bf Proof of Theorem \ref{thm:one}.} We first establish the equivalence of claims 2 and 3. If claim 2 holds, then
 claim 3 holds trivially. Suppose claim 3 holds and pick $\ell \in \{1,\ldots,m\}$. Let  $\theta_{v_1},\ldots,\theta_{v_m},\theta_{h_1},\ldots,\theta_{h_n}$, $\theta_{ij}$, $i=1,\ldots,m$, $j=1,\ldots,n$, denote the components of $\bm{\theta}_1$ and let $\tilde{\theta}_{v_1},\ldots,\tilde{\theta}_{v_m},\tilde{\theta}_{h_1},\ldots,\tilde{\theta}_{h_n}$, $\tilde{\theta}_{ij}$, $i=1,\ldots,m$, $j=1,\ldots,n$, denote the analogous components of $\bm{\theta}_2$. Then, by assumption,
\begin{align*}
\exp\left[ 2\theta_{v_\ell} +2 \sum_{j=1}^n  h_j \theta_{\ell j} \right] &= \frac{P_1\left(\{V_\ell = 1\} \cap \bigcap_{i=1 \atop i \neq \ell}^m \{V_i=v_i\} \cap \bigcap_{j=1}^n \{H_j=h_j\}\right)} {P_1\left(\{V_\ell = -1\} \cap \bigcap_{i=1 \atop i \neq \ell}^m \{V_i=v_i\} \cap \bigcap_{j=1}^n \{H_j=h_j\}\right)}\\
&= \frac{P_2\left(\{V_\ell = 1\} \cap \bigcap_{i=1 \atop i \neq \ell}^m \{V_i=v_i\} \cap \bigcap_{j=1}^n \{H_j=h_j\}\right)}
{P_2\left(\{V_\ell = -1\} \cap \bigcap_{i=1 \atop i \neq \ell}^m \{V_i=v_i\} \cap \bigcap_{j=1}^n \{H_j=h_j\}\right)}\exp\left[ 2\tilde{\theta}_{v_\ell} +2 \sum_{j=1}^n  h_j \tilde{\theta}_{\ell j} \right]
\end{align*}
for any $\{v_i\}_{i=1,i\neq \ell}^m,\{h_j\}_{j=1}^n \subset\{\pm 1\}$. Equating exponents, we have $\theta_{v_\ell} - \tilde{\theta}_{v_\ell} = \sum_{j=1}^n  h_j (  \tilde{\theta}_{\ell j} -\theta_{\ell j})$ for any $\{h_j\}_{j=1}^n \subset \{\pm 1\}$. Taking all $h_j=1$ or all $h_j=-1$ for $1 \leq j \leq n$ yields $\theta_{v_\ell} = \tilde{\theta}_{v_\ell}$. This result further implies $\tilde{\theta}_{\ell 1} -\theta_{\ell 1} = -  \sum_{j=2}^n  h_j (  \tilde{\theta}_{\ell j} -\theta_{\ell j})$ for
any given $\{h_j\}_{j=2}^n \subset \{\pm 1\}$, so that $\theta_{\ell 1}=\tilde{\theta}_{\ell 1}$ must hold. Iterating this argument, sequentially for each $j=2,\ldots,m$, shows $\theta_{\ell j}=\tilde{\theta}_{\ell j}$ for any $1 \leq j \leq n$ in addition to $\theta_{v_\ell}=\tilde{\theta}_{v_\ell}$.  As $\ell\in\{1,\ldots,m\}$ was arbitrary, we have $\theta_{v_i}=\tilde{\theta}_{v_i}$ and $\theta_{i j}=\tilde{\theta}_{i j}$ for any $1 \leq i \leq m$, $1 \leq j \leq n$.

@@

{\bf Proof.} The case $m=n=1$ follows from Lemma \ref{lemma:two}. We now use an induction argument, assuming that the result of Lemma \ref{lemma:one} holds for some given order of $(m,n)$ with $m,n\geq 1$ and show the the result continues to hold for $(m+1,n)$ or $(m,n+1)$. We shall treat the case $(m+1,n)$ (where the other case $(m,n+1)$ follows by symmetrical arguments).

@@

{\bf Proof.} Denote the four probabilities by which the random vector $(X,Y)$ assume pairs $(b,b),(b,a),(a,b),(a,a)$, respectively, as $c_1,c_2,c_3,1-c_1-c_2-c_3$ under $P_1$ and $d_1,d_2,d_3,1-d_1-d_2-d_3$ under $P_2$. The conditions $\E_1 X = \E_2 X$, $\E_1 Y = \E_2 Y$, $\E_1 X Y = \E_2 X Y$ with $b\neq a$ imply that
$$
(c_1+c_2) = (d_1+d_2), \quad  (c_1+c_3) = (d_1+d_3) \quad (b-a) c_1 + a (2c_1+c_2+c_3) =  (b-a) d_1 + a (2d_1+d_2+d_3).
$$
As $(2c_1+c_2+c_3) = (2d_1+d_2+d_3)$ and $b\neq a$, we conclude that $c_1=d_1$, from which it follows that $c_2=d_2$ and $c_3=d_3$. Hence, claim 1 implies claim 2 in Lemma \ref{lemma:two}. Claim 2 also trivially implies claim 1. $\Box$

@@

{\bf Proof of Theorem \ref{thm:two}.}  We first establish the equivalence of claims 2 and 3. If claim 2 holds, then claim 3 follows easily from the resulting uniform cell probabilities: $P(V_1= v_1,\ldots,V_m= v_m, H_1= h_1,\ldots,H_n=h_n|\bm{\theta})=2^{-mn}$ for any $v_i,h_j\in\{\pm 1\}$, $i=1,\ldots,m$, $j=1,\ldots,n$.

>>>>

---
title: "Supplement to: On the propriety of restricted Boltzmann machines"
author: "Andee Kaplan, Daniel Nordman, and Stephen Vardeman"
date: "November, 2016"
linestretch: 2 
output: 
  pdf_document:
    fig_caption: yes
    includes:
      in_header: ../latex/header.tex
    number_sections: yes
---

# On the one-to-one correspondence between parameters and moments in the RBM distribution

@@

**Proof of Theorem \ref{thm:one}.** We first establish the equivalence of claims 2 and 3. If claim 2 holds, then
 claim 3 holds trivially. Suppose claim 3 holds and pick $\ell \in \{1,\ldots,m\}$. Let  $\theta_{v_1},\ldots,\theta_{v_m},\theta_{h_1},\ldots,\theta_{h_n}$, $\theta_{ij}$, $i=1,\ldots,m$, $j=1,\ldots,n$, denote the components of $\bm{\theta}_1$ and let $\tilde{\theta}_{v_1},\ldots,\tilde{\theta}_{v_m},\tilde{\theta}_{h_1},\ldots,\tilde{\theta}_{h_n}$, $\tilde{\theta}_{ij}$, $i=1,\ldots,m$, $j=1,\ldots,n$, denote the analogous components of $\bm{\theta}_2$. Then, by assumption,
\begin{align*}
\exp\left[ 2\theta_{v_\ell} +2 \sum_{j=1}^n  h_j \theta_{\ell j} \right] &= \frac{P_1\left(\{V_\ell = 1\} \cap \bigcap_{i=1 \atop i \neq \ell}^m \{V_i=v_i\} \cap \bigcap_{j=1}^n \{H_j=h_j\}\right)} {P_1\left(\{V_\ell = -1\} \cap \bigcap_{i=1 \atop i \neq \ell}^m \{V_i=v_i\} \cap \bigcap_{j=1}^n \{H_j=h_j\}\right)}\\
&= \frac{P_2\left(\{V_\ell = 1\} \cap \bigcap_{i=1 \atop i \neq \ell}^m \{V_i=v_i\} \cap \bigcap_{j=1}^n \{H_j=h_j\}\right)}
{P_2\left(\{V_\ell = -1\} \cap \bigcap_{i=1 \atop i \neq \ell}^m \{V_i=v_i\} \cap \bigcap_{j=1}^n \{H_j=h_j\}\right)}\exp\left[ 2\tilde{\theta}_{v_\ell} +2 \sum_{j=1}^n  h_j \tilde{\theta}_{\ell j} \right]
\end{align*}
for any $\{v_i\}_{i=1,i\neq \ell}^m,\{h_j\}_{j=1}^n \subset\{\pm 1\}$. Equating exponents, we have $\theta_{v_\ell} - \tilde{\theta}_{v_\ell} = \sum_{j=1}^n  h_j (  \tilde{\theta}_{\ell j} -\theta_{\ell j})$ for any $\{h_j\}_{j=1}^n \subset \{\pm 1\}$. Taking all $h_j=1$ or all $h_j=-1$ for $1 \leq j \leq n$ yields $\theta_{v_\ell} = \tilde{\theta}_{v_\ell}$. This result further implies $\tilde{\theta}_{\ell 1} -\theta_{\ell 1} = -  \sum_{j=2}^n  h_j (  \tilde{\theta}_{\ell j} -\theta_{\ell j})$ for
any given $\{h_j\}_{j=2}^n \subset \{\pm 1\}$, so that $\theta_{\ell 1}=\tilde{\theta}_{\ell 1}$ must hold. Iterating this argument, sequentially for each $j=2,\ldots,m$, shows $\theta_{\ell j}=\tilde{\theta}_{\ell j}$ for any $1 \leq j \leq n$ in addition to $\theta_{v_\ell}=\tilde{\theta}_{v_\ell}$.  As $\ell\in\{1,\ldots,m\}$ was arbitrary, we have $\theta_{v_i}=\tilde{\theta}_{v_i}$ and $\theta_{i j}=\tilde{\theta}_{i j}$ for any $1 \leq i \leq m$, $1 \leq j \leq n$.

@@

**Proof.** The case $m=n=1$ follows from Lemma \ref{lemma:two}. We now use an induction argument, assuming that the result of Lemma \ref{lemma:one} holds for some given order of $(m,n)$ with $m,n\geq 1$ and show the the result continues to hold for $(m+1,n)$ or $(m,n+1)$. We shall treat the case $(m+1,n)$ (where the other case $(m,n+1)$ follows by symmetrical arguments).

@@

**Proof.** Denote the four probabilities by which the random vector $(X,Y)$ assume pairs $(b,b),(b,a),(a,b),(a,a)$, respectively, as $c_1,c_2,c_3,1-c_1-c_2-c_3$ under $P_1$ and $d_1,d_2,d_3,1-d_1-d_2-d_3$ under $P_2$. The conditions $\E_1 X = \E_2 X$, $\E_1 Y = \E_2 Y$, $\E_1 X Y = \E_2 X Y$ with $b\neq a$ imply that
$$
(c_1+c_2) = (d_1+d_2), \quad  (c_1+c_3) = (d_1+d_3) \quad (b-a) c_1 + a (2c_1+c_2+c_3) =  (b-a) d_1 + a (2d_1+d_2+d_3).
$$
As $(2c_1+c_2+c_3) = (2d_1+d_2+d_3)$ and $b\neq a$, we conclude that $c_1=d_1$, from which it follows that $c_2=d_2$ and $c_3=d_3$. Hence, claim 1 implies claim 2 in Lemma \ref{lemma:two}. Claim 2 also trivially implies claim 1. $\Box$

@@

**Proof of Theorem \ref{thm:two}.**  We first establish the equivalence of claims 2 and 3. If claim 2 holds, then claim 3 follows easily from the resulting uniform cell probabilities: $P(V_1= v_1,\ldots,V_m= v_m, H_1= h_1,\ldots,H_n=h_n|\bm{\theta})=2^{-mn}$ for any $v_i,h_j\in\{\pm 1\}$, $i=1,\ldots,m$, $j=1,\ldots,n$.
